Abstract Binding Trees

### 2015-10-21 paper.tex
wordsmithing

\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \IMode{#3}}

@@

Let $\Symbols$ be the infinite set of symbols; let $\Finsets$ be the free
cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
lift to renamings of operators' parameters; via the Grothendieck construction,
we can also consider the set $\int\Operators$ of operators
$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(U,\varrho)}$.

@@

The judgment $\IsOperator{U}{\vartheta}{\varrho}$ supports the structural
principles of weakening and exchange, because of the functoriality of
$\Operators$.

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$, about whose operators we may assert the following:

\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp})
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

However, consider the extension of the calculus with assignables. Then, we
shall have the following:

@@

Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
renaming $u\mapsto v$ will take $\mathsf{get}\{u\}$ to $\mathsf{get}\{v\}$.

@@

In general, we will have three kinds of context: metacontexts, variable
contexts, and parameter contexts. A metacontext $\Omega$ consists of bindings
of valences to metavariables; a variable context $\Gamma$ is a collection of
bindings of sorts to variables, and a parameter context $\Upsilon$ is a
collection of bindings of sorts to symbols.

@@

Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$.

>>>>

\newcommand\IsOperator[3]{\IMode{#1}\vdash \IMode{#2} : \OMode{#3}}

@@

Let $\Symbols$ be an infinite set of symbols; let $\Finsets$ be the free
cocartesian category over $\Symbols$. Then, fix a covariant presheaf of operators
$\Operators:\Finsets\times\Arities\to\Sets$ such that the arrows in $\Finsets$
lift to renamings of operators' parameters; via the Grothendieck construction,
we can also consider the set $\int\Operators$ of operators
$(U,\varrho,\vartheta)$ for $\Member{\vartheta}{\Operators(U,\varrho)}$.

@@

The judgment $\IsOperator{U}{\vartheta}{\varrho}$ enjoys the structural
properties of weakening and exchange via the functoriality of
$\Operators$.

@@

Operators are defined by specifying the fibres of $\int\Operators$ in which
they reside. For instance, consider the lambda calculus with a single sort,
$\SortExp$; we give its signature by asserting the following about its
operators:
\begin{gather*}
  \IsOperator{U}{\lambda}{
    \MkArity{\MkValence{\cdot}{\SortExp}{\SortExp}}{\SortExp})
  }\\
  \IsOperator{U}{\mathsf{ap}}{
    \MkArity{
      \MkValence{\cdot}{\cdot}{\SortExp},
      \MkValence{\cdot}{\cdot}{\SortExp}
    }{\SortExp}
  }
\end{gather*}

So far, we have made no use of symbols and parameters; however, consider the
extension of the calculus with assignables (references):

@@

Declaring a new assignable consists in providing an initial value, and an
expression with a free symbol (which shall represent the assignable in scope).
Weakening can be seen as inducing a ``degeneracy map'' on operators, whereas a
renaming $u\mapsto v$ will take $\mathsf{get}\{u\}$ to $\mathsf{get}\{v\}$.

@@

In general, we will have three kinds of context: metavariable contexts,
variable contexts, and symbol (parameter) contexts. A metavariable context
$\Omega$ consists of bindings of valences to metavariables; a variable context
$\Gamma$ is a collection of bindings of sorts to variables, and a parameter
context $\Upsilon$ is a collection of bindings of sorts to symbols.

@@

Let the judgment $\IsAbt{\Omega}{\Upsilon}{\Gamma}{M}{s}$ presuppose
$\IsMetaCtx{\Omega}$, $\IsSymCtx\Upsilon$, $\IsVarCtx\Gamma$ and $\IsSort{s}$,
meaning that $M$ is an abstract binding tree of sort $s$, with metavariables in
$\Omega$, parameters in $\Upsilon$, and variables in $\Gamma$. Then, the syntax
of abstract binding trees is inductively defined in three rules:
